The ITHIM model uses the comparative risk assesment framework to
estimate the population attributable fraction, $\af$.  In our case the
$\af$ represents the proportional change in national disease burden if
age and sex-specific active transport time means were changed.

\begin{equation}
\af = \frac{\int \! R(x)P(x) \, \mathrm{d}x  - \int \! R(x)Q(x) \, \mathrm{d}x}{\int \! R(x)P(x) \, \mathrm{d}x}
\end{equation}

Here $R(x)$ represents the risk when the individual has exposure $x$,
physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
density of individuals with exposure $x$, in the baseline and
alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of two independent random variables, travel-related
exposure and non-travel-related exposure.  Non-travel-related exposure
is modeled with a \logNormal{} distribution with the age-sex specific
mean ratios for non-travel-related exposure are estimated from the
ATUS \cite{ATUS}.  The referent group mean is treated as a variable
and the coefficient of variation is treated as constant across age-sex
strata and estimated with ATUS.  The distribution for travel-related
exposure is estimated from the time spent walking or cycling, i.e.,
active transport time, which as in the original ITHIM model has
a \logNormal{} distribution with constant coefficient of variation
across age-sex strata.  Travel-related exposure is estimated from
active transport time using assumptions about how much physical
activity is required for cycling and walking.

%does this involve using the CDC Wonder data? we could add a bit here about it, but if this is just national estimtes and shifts perhaps we don't need it. 

For the purposes of this implementation we limited our analyses to metropolitan regions of the United States. To idenfity urban populations we used the National Center for Health Statistics 2013 Urban-Rural Classification Scheme for Counties \cite{ingram2014}. We eliminated counties in the US with the classifcation of 'nonmetropolitan' (noncore or micropolitan). We used this and similar classification schemes as well as county and state identifiers in other datsets (NHTS, ATUS, CDC WONDER) to ensure consistency in our estimates for input parameters. 
