## ITHIM Package

### Install and Load ITHIM

```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
library("devtools")
install_github("syounkin/ITHIM", ref="devel")
library("ITHIM")
```

```{r misc, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
#install("~/ITHIM/")
#document("~/ITHIM/")
#install("~/ITHIM/")
fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

### Create the ITHIM object

#### Method 1: Using Input Files

We begin by creating the ITHIM object from which both the baseline and
scenarios are derived.  The necessary input parameters are contained
in 3 files, listed below.

```{r misc2, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = TRUE}
fileNamesList <- list(
    activeTransportTimeFile = "~/ITHIM/inst/activeTransportTime.csv",
    roadInjuriesFile = "~/ITHIM/inst/roadInjuries.csv",
    GBDFile = "~/ITHIM/inst/gbd.csv"
)

ITHIM <- createITHIM(fileNamesList)
```

Links to sample files can be found below.

* [activeTransportTime.csv](http://ithim.ghi.wisc.edu/activeTransportTime.csv)
* [roadInjuries.csv](http://ithim.ghi.wisc.edu/roadInjuries.csv)
* [gbd.csv](http://ithim.ghi.wisc.edu/gbd.csv)

These files can also be found in the devel branch of ITHIM repository.

* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/activeTransportTime.csv)
* [roadInjuries.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/roadInjuries.csv)
* [gbd.csv](https://github.com/syounkin/ITHIM/blob/devel/inst/gbd.csv)

