# Integrated Transport and Health Impacts Model

## R Package _ITHIM_
<!-- This document will eventually serve as a compendium to the R package
_ITHIM_.  It will describe how to use the package to implement the
ITHIM model, as well as a detailed description of the model itself.

As a working document it will serve as a place for SGY to post
results, questions, interpreations, notes, etc.-->

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
# install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r","devtools"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
```

```{r loadITHIM, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
document("~/ITHIM/")
install("~/ITHIM/")
#install_github("syounkin/ITHIM", ref="devel")
install_github("smbache/magrittr", ref="master")

library("magrittr")
library("ITHIM")

fig.width <- 6 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

## HHS Metro Region Parameter Sets

First we use NHTS to estimate mean walking and cycling times.

### NHTS

```{r compute, echo = FALSE, eval = TRUE}
compute <- TRUE
```

#### Create a handful of functions to help parse data

```{r parsely, eval = TRUE, echo = FALSE}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS01","HHS01","HHS01","HHS01","HHS01","HHS01","HHS02","HHS02","HHS03","HHS03","HHS03","HHS03","HHS03","HHS03","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS04","HHS05","HHS05","HHS05","HHS05","HHS05","HHS05","HHS06","HHS06","HHS06","HHS06","HHS06","HHS07","HHS07","HHS07","HHS07","HHS08","HHS08","HHS08","HHS08","HHS08","HHS08","HHS09","HHS09","HHS09","HHS09","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

countPeople <- function(data){
 nrow(data %>% group_by(HOUSEID, PERSONID) %>% dplyr::summarize(n()))
}

tripTime <- function(data, mode = "walk"){
 return(sum(subset(data, AT == mode)$TRVL_MIN, na.rm = TRUE))
}

reformatData <- function(data){
foo <- matrix(unlist(data), nrow = 7, ncol = 2)
foo <- rbind( c(0, 0), foo)
dimnames(foo) <- list(paste0("ageClass",1:8), c("F","M"))
return(foo[,c("M","F")])
}

```

#### Read Two of the Four NHTS Data Tables

```{r NHTS, eval = compute, echo = FALSE}
DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")

NHTS.trips <- with(DAYV2PUB,{
    data.frame(HOUSEID = HOUSEID,
                   PERSONID = PERSONID,
                   TDCASEID = TDCASEID,
                   AT = ifelse(TRPTRANS == "22", "cycle",
                        ifelse(TRPTRANS == "23", "walk",
                        ifelse(!is.na(TRPTRANS), "other", NA))),
                   TRVL_MIN = as.numeric(TRVL_MIN),
                   TRVLCMIN = ifelse(as.numeric(TRVLCMIN) != 0, as.numeric(TRVLCMIN), NA),
                   URBRUR  = as.factor(URBRUR),
                   MSACAT = as.factor(MSACAT),
                   MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                   TRPTRANS = as.factor(TRPTRANS),
                   TRPMILES = as.numeric(TRPMILES)
)})

PERV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/PERV2PUB.CSV", na.strings = "-9")
names(PERV2PUB) <- c("HOUSEID","PERSONID","VARSTRAT","WTPERFIN","SFWGT","HH_HISP","HH_RACE","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","WRKCOUNT","FLAG100","LIF_CYC","CNTTDTR","BORNINUS","CARRODE","CDIVMSAR","CENSUS_D","CENSUS_R","CONDNIGH","CONDPUB","CONDRIDE","CONDRIVE","CONDSPEC","CONDTAX","CONDTRAV","DELIVER","DIARY","DISTTOSC","DRIVER","DTACDT","DTCONJ","DTCOST","DTRAGE","DTRAN","DTWALK","EDUC","EVERDROV","FLEXTIME","FMSCSIZE","FRSTHM","FXDWKPL","GCDWORK","GRADE","GT1JBLWK","HHRESP","HHSTATE","HHSTFIPS","ISSUE","OCCAT","LSTTRDAY","MCUSED","MEDCOND","MEDCOND6","MOROFTEN","MSACAT","MSASIZE","NBIKETRP","NWALKTRP","OUTCNTRY","OUTOFTWN","PAYPROF","PRMACT","PROXY","PTUSED","PURCHASE","R_AGE","R_RELAT","R_SEX","RAIL","SAMEPLC","SCHCARE","SCHCRIM","SCHDIST","SCHSPD","SCHTRAF","SCHTRN1","SCHTRN2","SCHTYP","SCHWTHR","SELF_EMP","TIMETOSC","TIMETOWK","TOSCSIZE","TRAVDAY","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","WEBUSE","WKFMHMXX","WKFTPT","WKRMHM","WKSTFIPS","WORKER","WRKTIME","WRKTRANS","YEARMILE","YRMLCAP","YRTOUS","DISTTOWK","TDAYDATE","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","HH_CBSA","HHC_MSA")

NHTS.person <- with(PERV2PUB,{
                    data.frame(HOUSEID = HOUSEID,
                               PERSONID = PERSONID,
                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
                               AGE = as.factor(convertToAgeClass(R_AGE)),
                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                               HHSTATE = as.character(HHSTATE),
                               HHS = as.character(convertToHHS(as.character(HHSTATE))), stringsAsFactors = FALSE

)
})

```

```{r summary, fig.width = 20, fig.height = 10, eval = compute, echo = FALSE, message = FALSE, warning = FALSE}
walkHours <- with(subset(NHTS.trips, {
    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
}), TRVL_MIN/60)

walkMiles <- with(subset(NHTS.trips, {
    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVL_MIN*60 < 8 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 0.5
}), TRPMILES)

cycleHours <- with(subset(NHTS.trips, {
      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVL_MIN*60 < 25 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 1
}), TRVL_MIN/60)

cycleMiles <- with(subset(NHTS.trips, {
      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVL_MIN*60 < 25 & TRVL_MIN > 5 #& TRPMILES/TRVL_MIN*60 > 1
}), TRPMILES)

walkSpeed <- median(walkMiles/walkHours, na.rm = TRUE)
cycleSpeed <- median(cycleMiles/cycleHours, na.rm = TRUE)

Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, Speed = walkMiles/walkHours, mode = "walk")
Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, Speed = cycleMiles/cycleHours, mode = "cycle")

D <- rbind(Dwalk,Dcycle)

Dspeed <- data.frame(speed = c(1,1), intercept = log10(c(walkSpeed,cycleSpeed)), mode = c("walk","cycle"))

g <- ggplot(D, aes(Hours, Miles, colour = mode)) + geom_point()
g <- g + geom_abline(aes(slope = speed, intercept = intercept, colour = mode), Dspeed) + facet_grid(. ~ mode)
g + scale_y_log10() + scale_x_log10() + theme_bw()

g <- ggplot(D, aes(Speed, ..density.., fill = mode)) + geom_histogram() + theme_bw() + scale_x_continuous(name="Speed") + scale_y_continuous(name="Density") + facet_grid(. ~ mode, scales = "free")
g

```

#### Subset and Merge

The PERSON table contains the family members of the person surveyed,
right?  We want only the peole who were surveyed for their travel
times.  So, we can\'t use everyone in the PERSON table.  If someone
had zero trip time they would not exist in the TRIP table.  So, if we
use just the TRIP table we miss people with zero trip time. So how do
we merge?  We use an inner join and assume that their is a negligible
number of people with zero trip time.  Not very many people never stay
in the house all day.  Please correct me if I am mistaken abiout the
nature of the PERSON or TRIP tables.

```{r subset, eval = compute, fig.width = 10, echo = FALSE}
NHTS.person <- subset( NHTS.person, MSA == TRUE )
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
saveRDS(NHTS, file = "./data/NHTS.rds")
```

```{r loadNHTS, eval = !compute}
NHTS <- readRDS(file = "./data/NHTS.rds")
```

#### Compute National Metro Average

```{r natl, eval = TRUE, echo = FALSE, warning = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$SEX),function(x) split(x, x$AGE))
counts <- lapply(listOfLists, function(x) lapply(x, countPeople))
walkLength <- lapply(listOfLists, function(x) lapply( x, tripTime, mode = "walk"))
cycleLength <- lapply(listOfLists, function(x) lapply( x, tripTime, mode = "cycle"))

countMatrix <- reformatData(counts)
walkMatrix <- reformatData(walkLength)
cycleMatrix <- reformatData(cycleLength)

meanWalkList <- walkMatrix/countMatrix*7
meanCycleList <- cycleMatrix/countMatrix*7

RwtList <- ifelse(meanWalkList == 0, 1e-4, meanWalkList)
RctList <- ifelse(meanCycleList == 0, 1e-4, meanCycleList)

RwtNational <- RwtList/RwtList[3,2]
RctNational <- RctList/RctList[3,2]
muwtNational <- meanWalkList[3,2]
muctNational <- meanCycleList[3,2]

rm("listOfLists","counts","walkLength","cycleLength","countMatrix","walkMatrix","cycleMatrix","meanWalkList","meanCycleList","RwtList","RctList","muwtList","muctList")
```

#### Compute Mean Trip Lengths

```{r count, eval = TRUE, echo = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$HHSTATE),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)
meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) ifelse(y == 0, 0, x/y*7), SIMPLIFY = FALSE)

zeroReferentCase <- function(x){
  if(x[3,2]==0){
    return(matrix(1e-3,nrow=nrow(x),ncol=ncol(x), dimnames = dimnames(x)))
  }else{
    return(ifelse(x==0, 1e-3, x/x[3,2]))
  }
}

RwtList <- lapply( meanWalkList, zeroReferentCase)
RctList <- lapply( meanCycleList, zeroReferentCase)
muwtList <- lapply( meanWalkList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
muctList <- lapply( meanCycleList, function(x) ifelse(x[3,2]==0,1e-3, x[3,2]))
```


## Create List of ITHIM Objects


```{r foobar, eval = TRUE, echo = FALSE}
sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
    return(sum(unlist(data.frame(compareModels(ITHIM.baseline,ITHIM.scenario)$daly.delta)[-1,]))) # AgeClass 1 is removed from totals
    }
```

```{r ITHIMObjs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
stateVec <- c("CA","TX","NY","FL") #c("CA","TX","NY","FL","VA","NC","AZ")
stateVecOrdered <- c("TX","FL","NY","CA") #c("NC","TX","VA","FL","AZ","NY","CA")

upper <- 10; length <- 5; wd <- upper/length;

DALYsTotal <- data.frame()
ITHIMList <- list()
ITHIM.default <- createITHIM()
ITHIM.default <- updateITHIM(ITHIM.default, "cvNonTravel", 3)
ITHIM.default <- updateITHIM(ITHIM.default, "cv", 1.65)
ITHIM.default <- updateITHIM(ITHIM.default, "muNonTravel", 4)
ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)

ITHIM.national <- ITHIM.default
ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)

for (HHSTATE in stateVec){
    ITHIMList[[HHSTATE]] <- ITHIM.default
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
  }

for( muNonTravel in seq(0, upper, by = wd)){

  rm(DALYs)

  for (HHSTATE in stateVec){
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
  }

  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)

  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
  DALYsTotal <- rbind(DALYsTotal, DALYs)
}

DALYsTotal <- within(DALYsTotal, {
  State <- factor(State, levels = stateVecOrdered)
})

p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))

p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper))

p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point()

```

Do it again, but with quintiles

```{r ITHIMObjsQuintiles, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, echo = TRUE, warning = FALSE}
DALYsTotal2 <- data.frame()
ITHIMList <- list()

ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", seq(1, 9, by = 2)/10 )

ITHIM.national <- ITHIM.default
ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)

for (HHSTATE in stateVec){
    ITHIMList[[HHSTATE]] <- ITHIM.default
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muwt", muwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muct", muctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rwt", RwtList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "Rct", RctList[[HHSTATE]])
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "GBD", ITHIM.national$parameters$GBD)
  }

for( muNonTravel in seq(0, upper, by = wd)){

  rm(DALYs)

  for (HHSTATE in stateVec){
    ITHIMList[[HHSTATE]] <- updateITHIM(ITHIMList[[HHSTATE]], "muNonTravel", muNonTravel)
  }

  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)

  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
  DALYs <- data.frame(State = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
  DALYsTotal2 <- rbind(DALYsTotal2, DALYs)
}

DALYsTotal2 <- within(DALYsTotal2, {
  State <- factor(State, levels = stateVecOrdered)
})

DALYsTotal <- rbind(data.frame(DALYsTotal,quints="Percentiles"), data.frame(DALYsTotal2,quints="Quintiles"))

p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY/1e3, colour = State))

p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( . ~ quints)

p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( . ~ quints)

```


### Visualization of trip time matrices

```{r meanMatrices, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}

ITHIMList <- c(ITHIMList, list("Natl." = ITHIM.national))

meanMatrix <- function(ITHIM, mode = "Walking"){
   if(mode == "Walking"){
      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
   }else if (mode == "Cycling"){
      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
   }else{
      message("error in meanMatrix function")
   }
}

meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "Walking")
meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "Cycling")

D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))
names(D) <- c("ageClass", "mode", "Sex", "time", "State")

D <- within(D,{
  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
  Sex <- factor(Sex, levels = c("F","M"))
})

p <- ggplot(subset(D,ageClass != 1), aes(x = ageClass, y = time, fill = Sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Travel Time per Week (minutes)") + facet_grid( mode ~ State, scales = "free_y") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 5))

```

```{r California, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
ITHIM.national <- updateITHIM(ITHIM.national, parName = "muNonTravel", 10)

ITHIM.CA <- ITHIMList[["CA"]]
ITHIM.CA <- updateITHIM(ITHIM.CA, parName = "muNonTravel", 10)
foo1 <- compareModels(ITHIM.national,ITHIM.CA)
foobar1 <- data.frame(cbind(melt(foo1$daly.delta),paste0("ageClass",1:8)),State = "CA")

foobar <- foobar1
names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")

## ITHIM.NY <- ITHIMList[["NY"]]
## ITHIM.NY <- updateITHIM(ITHIM.NY, parName = "muNonTravel", 10)
## foo2 <- compareModels(ITHIM.national,ITHIM.NY)
## foobar2 <- data.frame(cbind(melt(foo2$daly.delta),paste0("ageClass",1:8)),State = "NY")
## foobar <- rbind(foobar1,foobar2)
## names(foobar) <- c("delta.DALY","Sex","Disease","ageClass","State")

foobar <- within(foobar, {
  Disease[Disease == "BreastCancer"] <- "Breast Cancer"
  Disease[Disease == "ColonCancer"] <- "Colon Cancer"
  Disease <- factor(Disease, levels = c("Breast Cancer", "Colon Cancer", "Dementia", "Depression", "Diabetes"))
})

p <- ggplot( subset(foobar,ageClass != "ageClass1"), aes(ageClass,-delta.DALY/1e3, fill = Sex) )

p + geom_bar(stat="identity", position = "dodge") +
theme(axis.title=element_blank()) +
facet_grid( Disease ~ ., scales = "free") +
scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") +
scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+"))

```

```{r counts, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = phi, warning = FALSE, message = FALSE, echo = FALSE}
D <- melt(counts)
names(D) <- c("count","AgeClass","Sex", "State")
D <- within(D,{
  State <- factor(State, levels = c("NC","TX","VA","FL","AZ","NY","CA","Natl."))
  Sex <- factor(Sex, levels = c("F","M"))
})
p <- ggplot(subset(D,AgeClass != "00-04"), aes(AgeClass, count/1e3))
p + geom_bar(stat="identity", position = "dodge") + scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Individuals Sampled (Thousands)") + facet_grid( State ~ .) + theme_bw() + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=6))
```


## Figure 1

```{r methodsPaper, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
ITHIM.default <- updateITHIM(ITHIM.default, "quantiles", (1:99)/100)

ITHIM.national <- ITHIM.default
ITHIM.national <- updateITHIM(ITHIM.national, "muwt", muwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "muct", muctNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rwt", RwtNational)
ITHIM.national <- updateITHIM(ITHIM.national, "Rct", RctNational)


ITHIM.scenario <- ITHIM.national
ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=1)/10)
ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=1)/10)

pWalk <- ITHIM.national$means$pWalk[3,2]

DALYs <- c(); DALYs.quints <- c();
atVec <- seq(45, 120, length.out = 5)
muNonTravelVec <- c(1, 10, 50, 100, 150, 200)
DTotal <- data.frame()

for(at in atVec ){  # minutes active transport/week
  for( muNonTravel in muNonTravelVec ){

    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muwt", at*pWalk)
    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muct", at*(1-pWalk))
    ITHIM.scenario <- updateITHIM(ITHIM.scenario, "muNonTravel", muNonTravel)

    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muwt", at*pWalk)
    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muct", at*(1-pWalk))
    ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario.quints, "muNonTravel", muNonTravel)

    ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)

    DALYs <- sumDALY(ITHIM.national, ITHIM.scenario)
    DALYs.quints <- sumDALY(ITHIM.national.quints, ITHIM.scenario.quints)

    D <- data.frame(quants = c("percentiles","quintiles"), DALYs = c(DALYs,DALYs.quints),at = at, muNonTravel = muNonTravel)

    DTotal <- rbind(D,DTotal)

  }
}

p <- ggplot(DTotal, aes(x = at,y = -DALYs/1e3, colour = quants))
p + stat_smooth( method = loess) + geom_point() + theme_bw() + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + scale_x_continuous(name="Active Transport (minutes/week) (Referent Group)") + facet_grid( . ~ muNonTravel) + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 8))

```

## Figure 2

```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
D <- rbind(getWalkTime(ITHIM.default),getWalkSpeed(ITHIM.default),getCycleTime(ITHIM.default),getNonTravelMETs(ITHIM.default))
D <- within(D, sex <- factor(sex, levels = c("F","M")))

p <- ggplot(D, aes(x = ageClass, y = mu, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Travel Time per Week (minutes)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ ., scales = "free_y")
```

```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
D <- melt(ITHIM.default$parameters$GBD)
names(D) <- c("disease","sex","ageClass","type","burden", "sex2", "disease2")
D <- within(D, sex <- factor(sex, levels = c("F","M")))

p <- ggplot(subset(D,disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes")), aes(x = ageClass, y = burden, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ disease, scales = "free_y")
```
