## ITHIM Tutorial

### Installing the package

```{r figwidth, eval = TRUE,  warning = FALSE, message = FALSE, echo = FALSE}
fig.width <- 6
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

First load the package from Github.  To do this you need the package
devtools.

```{r, libs, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
options(repos='http://cran.rstudio.com/')
install.packages("devtools")
```

Now load the package and install ITHIM.

```{r libs2, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
library("devtools")
install_github(repo = "syounkin/ITHIM", ref = devel)
library("ITHIM")
```

### Built-in Example

```{r example, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
example("ITHIM-package")
```

### Step-by-step

Now that ITHIM is installed and loaded the first step is to create the
baseline model.  To do so, using defaults built into the package, use
createITHIM as follows.

```{r create, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
ITHIM.baseline <- createITHIM()
```

There are currently fourteen parameters, many of which are matrices,
in the model.  (Recall that we are only using the physical activity
component.)

```{r parameters, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
names(ITHIM.baseline$parameters)
```

We have means and mean matrices for walk time, cycle time and
non-travel-related activity and coefficients of variation for active
transport time and non-travel-related activity.  nAgeClass has always
been 8 and meanType is always "referent".  GBD is an array of disease
burden estimates and quantiles is the vector of quantiles used for the
approximation of the PAF integral.

Now to create the scenario we need only change one parameter.
Parameters may be changed with updateITHIM.  First we change the mean
walking time muwt to 120 minutes.

```{r scenario, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
ITHIM.scenario <- updateITHIM(ITHIM.baseline, "muwt", 120)
```
Now to run the comparative risk assesment use compareModels.
```{r CRA, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
comparativeRisk <- compareModels(ITHIM.baseline, ITHIM.scenario)
```

```{r AF, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = TRUE}
comparativeRisk$AF$ColonCancer
```
